---
title: "Supporting Information"
output:
  html_document:
    df_print: paged
---

##1.0 Explanation of Methods
To begin to explore the use and interdisciplinary nature of the *hydrologic connectivity* concept, we conducted a cursory literature review using R statistical software (R Core Team, 2018), the bibliometrix package (Aria and Cuccurullo, 2017), and methods similar to Bernhardt et al., (2017). To do this, we queried the ISI Web of Knowledge database for articles that used the term "hydrologic connectivity" between 1998 and 2018. We then discretized those articles by journal discipline; classified keywords by physical, chemical, or biological function; and finally, identified keywords that referred to specific hydrologic units. Below, we provide examples tables that demonstrate how journals and key terms were classified. Further, we provide the Rcode used to create Figure 1.

```{r setup, echo=FALSE, results='hide', message=FALSE}
#Define relevant libraries
library(tidyverse)
library(bibliometrix)
library(kableExtra)
library(stringr)

#Define working dir
working_dir<-"/nfs/njones-data/Research Projects/JAWRA_Special_Issue/lit_analysis/"

#import .bib file
bib<-readFiles(paste0(working_dir,"connectivity_lit.bib"))
bib<-bibliometrix::convert2df(bib, dbsource = 'isi', format='bibtex')
results <-bibliometrix::biblioAnalysis(bib, sep = ";")
results <- summary(object = results, k=500, pause = F)

#Read jounral
journals<-readxl::read_excel(paste0(working_dir,"classification.xlsx"), sheet="journals")
key_words<-readxl::read_excel(paste0(working_dir,"classification.xlsx"), sheet="key_words")
hydro_units<-readxl::read_excel(paste0(working_dir,"classification.xlsx"), sheet="hydro_units")
```

##2.0 Example Journal Classification
We classified journals into 5 categories: hydrology, ecology, biogeochemistry, environmental science, and general earth science.  All journals were assigned to a single category. Note, categories are similar to those used in Bernhardt et al., (2017). See Table 1 for examples of how journals were categorized.

```{r journal, echo=FALSE, message=FALSE,paged.print=TRUE}
#organize data
df<-as_tibble(results$MostRelSources) %>%
  #Rename collumns
  rename(journal = `Sources       `, pubs=Articles) %>%
  #format collumns
  mutate(#trim white space in journal names
         journal = paste(trimws(journal)), 
         #make pubs numeric
         pubs = as.numeric(paste(pubs))
         ) %>%
  #Remove journals with <4 citations
  filter(pubs>3) %>%
  #left join with journal classification table
  left_join(.,journals) %>%
  #redo journal names
  mutate(journal=str_to_title(journal))

#Change Colnames
colnames(df)<-c("Journal", "Count", "Disciplinary Category")

#Print top 10 
kable(df[1:10,], 
      "latex",
      booktabs=T,
      align = c("l", "c", "l"), 
      caption = "Separation of journals into distinct disciplinary categories") %>%
  kable_styling(latex_options = c("striped","hold_position"))
```

##3.0 Example Functional Classification
We identified key words that related to physical, chemical, and biological functions of aquatic systems from across the selected manuscripts. Individual keywords could only relate to one functional category, but each paper could contain key words from multiple functional categories. See Table 2 for examples of how key words were categorized.

```{r fun_cat, echo=FALSE, message=FALSE,paged.print=TRUE}
#Estimate number of pubs 
df<-key_words %>%
  #format key words tibble
  filter(str_detect(class, "function")) %>%  
  #select unique combinations of ID and class
  na.omit(.) 

#Change functional name for output
df$class[df$class=="biological function"]<-"Biological"
df$class[df$class=="chemical function"]<-"Chemical"
df$class[df$class=="physical function"]<-"Physical"

#rename collumns
colnames(df)<-c("Key work", "Count", "Function")

#Print top 10 
kable(df[1:10,], 
      "latex",
      booktabs=T,
      align = c("l", "c", "l"), 
      caption = "Classification of key words by function") %>%
  kable_styling(latex_options = c("striped","hold_position"))

```

##4.0 Example Hydrologic Unit Designation
We also identified key words that related to differnt hydrologic units. Individual keywords could only relate to one hydrologic unit, but each paper could be tied to multiple hydrologic units. See Table 3 for examples of how key words were categorized.
```{r hyd_cat, echo=FALSE, message=FALSE,paged.print=TRUE}
#Estimate number of pubs 
df<-tibble(key_word = bib$ID) %>%
  #create long format
  separate_rows(., key_word,sep=";") %>%
  #trim white space
  mutate(key_word = trimws(paste(key_word))) %>%
  #join hydro units classification table
  left_join(. , hydro_units) %>% 
  #remove NA
  na.omit() %>%
  #tally
  group_by(key_word, hydro_units) %>%
  tally() %>%
  arrange(desc(n))

#Change hydro units for display names
df$hydro_units[df$hydro_units=="Watershed"] <-"Watersheds and Stream Networks"
df$hydro_units[df$hydro_units=="River"] <-"Rivers and Streams"
df$hydro_units[df$hydro_units=="Wetland"] <-"Wetlands"
df$hydro_units[df$hydro_units=="Riparian"] <-"Riparian Zones and Floodplains"
df$hydro_units[df$hydro_units=="Hillslope"] <-"Critical Zone, Hillslopes, and Soils"
df$hydro_units[df$hydro_units=="Lake"] <-"Lakes, Reservoirs, and Ponds"
df$hydro_units[df$hydro_units=="Hyporheic Zone"] <-"Hyporheic Zones"

#Reorder collumns
df<-tibble(key_word = df$key_word, 
           Count    = df$n, 
           hydro    = df$hydro_units)

#rename collumns
colnames(df)<-c("Key work", "Count", "Hydrologic unit")

#Print top 10 
kable(df[1:10,], 
      "latex",
      booktabs=T,
      align = c("l", "c", "l"), 
      caption = "Classification of key words by hydrologic unit") %>%
  kable_styling(latex_options = c("striped","hold_position"))
```

##5.0 Rcode
```{r code, eval=F}
##################################################################################
#Step 1:  Setup Workspace---------------------------------------------------------
##################################################################################
#Clear Memory
rm(list=ls(all=TRUE))

#required packages
library(tidyverse)    #tidyverse family of packages
library(gridExtra)    #plotting
library(treemapify)   #plotting
library(packcircles)  #plotting
library(bibliometrix) #bibliography analysis
library(lubridate)    #date analysis
library(readxl)       #data import

#Define working directory
data_dir<- #Insert working directory here

##################################################################################
#Step 2:  Wrangle Citaiton Data---------------------------------------------------
##################################################################################
#Import .bib file info
#read .bib file
bib<-readFiles(paste0(data_dir,"connectivity_lit.bib"))

#Convert to df
bib<-convert2df(bib, dbsource = 'isi', format='bibtex')

#create summary
results <- biblioAnalysis(bib, sep = ";")
results <- summary(object = results, k=500, pause = F)

#read in manual classification tables (see tables 1-3 for examples)
journals<-read_excel(paste0(data_dir,"classification.xlsx"), sheet="journals")
key_words<-read_excel(paste0(data_dir,"classification.xlsx"), sheet="key_words")
hydro_units<-read_excel(paste0(data_dir,"classification.xlsx"), sheet="hydro_units")

##################################################################################
#Step 3:  Plotting----------------------------------------------------------------
##################################################################################
#A) Publication rate over time----------------------------------------------------
#create tibble of year and # of pubs
df<-as_tibble(results$AnnualProduction) 
colnames(df)<-c("year", "pubs")
df$year<-as.numeric(paste(df$year))

#create exponential model
model<-lm(log(pubs)~year, df)
model_data<-tibble(x=df$year, y=exp(fitted(model)))

#create plot
p1<-df %>%
  #Start plotting device
  ggplot(aes(x=year, y=pubs)) +
  #Add points
  geom_point(col="grey30") +
  #Add exp model
  geom_line(data = model_data, 
            aes(x, y), 
            col="grey70",
            size=1.3,
            linetype = 2) +
  #Add text describing model
  geom_text(aes(x=2003, y=75), label="RSQ=0.94\np<0.001", col="grey30")+
  #Axis labels
  labs(x="Year", y = "Number of Publications") +
  #Add Theme Elements
  theme_bw(base_size=12) +
  theme(
    #bold font for both axis text
    axis.text=element_text(face="bold"),
    #set thickness of axis ticks
    axis.ticks=element_line(size=0.4),
    #remove plot background
    plot.background=element_blank(), 
    #Decrease right margin for final plot
    plot.margin=unit(c(5.5, 12, 5.5, 5.5), "points")
  )

#B) Treemap of discpline and journal---------------------------------------------
#create tibble of journal and number of pubs
df<-as_tibble(results$MostRelSources) %>%
  #Rename collumns
  rename(journal = `Sources       `, pubs=Articles) %>%
  #format collumns
  mutate(#trim white space in journal names
         journal = paste(trimws(journal)), 
         #make pubs numeric
         pubs = as.numeric(paste(pubs))
         ) %>%
  #Remove journals with <4 citations
  filter(pubs>3) %>%
  #left join with journal classification table
  left_join(.,journals) %>%
  #redo journal names
  mutate(journal=str_to_title(journal))

#Create color pallete matrix
color<- tibble(classification = c("environmental science","hydrology",
                                  "ecology","biogeochemistry", "earth science"), 
             org=c(4,2,3,1,5))
df<-left_join(df, color) %>% 
  arrange(org) %>%
  mutate(color = factor(org))

#Plot tree map
p2<-ggplot(df, aes(area = pubs, label=journal, subgroup = classification, fill=color)) +
  geom_treemap() +
  geom_treemap_text(colour = "white", place = "centre",grow = F, reflow=T) +
  geom_treemap_subgroup_border(lwd=0.2, col="grey90") +
  scale_fill_brewer(palette="Set1") +
  theme_bw() +
  theme(legend.position="none")

#C) Circle plot of functions-----------------------------------------------------
#Add unique ID to bib tibble
bib$UID<-seq(1, nrow(bib))

#Create long-fromat tibble with ID and key word (for both author and ISI defined IDs)
w1<-tibble(ID = bib$UID, key_word = bib$ID) %>%
  #create long format
  separate_rows(., key_word,sep=";") %>%
  #trim white space
  mutate(key_word = trimws(paste(key_word)))
w2<-tibble(ID = bib$UID, key_word = bib$DE) %>%
  #create long format
  separate_rows(., key_word,sep=";") %>%
  #trim white space
  mutate(key_word = trimws(paste(key_word)))
w<-bind_rows(w1, w2)

#Estimate number of pubs 
df<-key_words %>%
  #format key words tibble
  select(key_word, class) %>% 
  filter(str_detect(class, "function")) %>%  
  #join tibbles [ie journal key words and key word classifications]
  left_join(w, .) %>% 
  #select unique combinations of ID and class
  na.omit(.) %>%
  select(ID, class) %>%
  distinct(ID, class) %>%
  #tally functions represented by different pubs
  count(class) 

#setup plotting layout
packing<-circleProgressiveLayout(df$n, sizetype = "area")
df <- cbind(df, packing)
df.gg<-circleLayoutVertices(packing, npoints = 50)
df.gg$value=rep(df$class, each=51)

#Format Names for text
df$class[df$class=="biological function"]<-"Biological"
df$class[df$class=="chemical function"]<-"Chemical"
df$class[df$class=="physical function"]<-"Physical"

#plot
p3<-ggplot() + 
  #Plot Polygongs
  geom_polygon(data = df.gg, 
               aes(x, y, group = id, fill=as.factor(id)), 
               colour = "black") +
  scale_fill_manual(values = c("#4daf4a","#e41a1c",  "#377eb8")) +
  #Add text
  geom_text(data = df, aes(x, y, label = paste0(class,"\n[n=",n,"]"), size=4)) +
  #theme 
  theme_void(base_size = 12) + 
  theme(legend.position="none") +
  coord_equal()

#D) Lolly-pop plot of hydrologic units------------------------------------------
#Create tibble of publication counts
df<-tibble(ID = bib$UID, key_word = bib$ID) %>%
  #create long format
  separate_rows(., key_word,sep=";") %>%
  #trim white space
  mutate(key_word = trimws(paste(key_word))) %>%
  #join hydro units classification table
  left_join(. , hydro_units) %>%
  #Remove irrelivant key words
  select(ID, hydro_units) %>%
  na.omit(.) %>%
  distinct(.) %>%
  #Tally hydro units
  count(hydro_units)

#preserve order in plot by changing hydro_units to factor
df<- df %>% 
  arrange(n) %>%
  mutate(hydro_units = factor(hydro_units, hydro_units))

#Create plot
p4<-ggplot(df, aes(x=hydro_units, y=n)) + 
  #Add lollipops
  geom_segment(aes(x=hydro_units, xend=hydro_units, y=0, yend=n), color="grey30") +
  geom_point(col="blue", size=4, alpha=0.8) +
  #Add lables
  labs(x="Hydrologic Units", y="Number of Publications")+
  #Add theme info
  theme_bw(base_size = 12) +
  coord_flip() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.y =  element_text(margin=margin(l=2), hjust=1)
  )
  
#Export multi plot --------------------------------------------------------
ggdraw()+
  draw_plot(p1, x=0.05,   y=0.5, width =0.55, height =0.5) +
  draw_plot(p3, x=0.6, y=0.5, width =0.4, height =0.5) +
  draw_plot(p4, x=0.05,   y=0,   width =0.9, height =0.5) +
  draw_plot_label(c("A)", "B)", "C)"), 
                  x=c(0,0.6, 0), 
                  y=c(1,1,0.5), 
                  size=12)
ggsave("figure1.pdf", device="pdf", width = 6, height= 4.5, units="in")
```

##6.0 Works Cited
Aria, M., C. Cuccurullo. 2017. "bibliometrix: An R-tool for comprehensive science mapping analysis." *Journal of Informatics* 11(4): 959-975. doi: 10.1016/j.joi.2017.08.007.

Bernhardt, E.S., J.R. Blaszczak, C.D. Ficken, M.L. Fork, K.E. Kaiser, and E.C. Seybold. 2017. “Control Points in Ecosystems: Moving Beyond the Hot Spot Hot Moment Concept.” *Ecosystems* 20(4): 665–682. doi: 10.1007/s10021-016-0103-y.

R Core Team. 2018. *R: A Language and Environment for Statistical Computing.* Vienna, Austria: R Foundation for Statistical Computing. http://www.R-project.org/.